Global Reach or Out of Reach

### 2019-05-13 Paper_markdown.Rmd
fix paths to datasets and .bib file; omit calls to `View()`; fix headings

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Global Reach or Out of Reach"
author: 
- name: Addison Huygens
  affiliation: University of Iowa
abstract: "Research which has examined the effects of naval capabilities on conflict behavior have done so by examining different measures for the size of a state's navy. However, this does not capture the actualy capabilities of a country's fleet. In this article I create a new measurement for naval capability and use it to test the effect of a hegemon's naval power on the conflict behavior of other states in the international system. I find that higher levels of naval power held by the hegemon reduces the liklihood that a militarized dispute will occur."
keywords: "Naval power hegemony"
date: "May 10, 2019"
fontsize: 12pt
spacing: single
bibliography: Naval.bib
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
---

@@

dyadic_mid_3_1_may_2018 <- as.data.frame( read_dta("C:/Users/packe/Documents/Methods_3/Paper/Datasets/dyadic_mid_3.1_may_2018.dta"))
View(dyadic_mid_3_1_may_2018)

@@

#Relative capabilities data
NMC_5_0 <- as.data.frame(read_dta("C:/Users/packe/Documents/Methods_3/Paper/Datasets/NMC_5_0.dta"))
View(NMC_5_0)
NMC <- select(NMC_5_0, ccode, year, irst, milex, milper, pec, tpop, upop, cinc)
NMCA <- NMC %>%
  rename(
    ccodea = ccode,
    irsta = irst,
    milexa = milex,
    milpera = milper,
    peca = pec,
    tpopa = tpop,
    upopa = upop,
    cinca = cinc
  ) 
NMCA

@@

#Importing naval data, removing all but the data for the U.S., keeping only relevant variables.
crisher_souva_navaldata_II <- as.data.frame(read_dta("C:/Users/packe/Documents/Methods_3/Paper/Datasets/crisher_souva_navaldata_II.dta"))
Naval1 <- filter(crisher_souva_navaldata_II, cowcode == 2)
Naval2 <- select(Naval1, year, totton, tonn_prop, milex, milper)
USNaval <- Naval2 %>%
  rename(
    ustoton = totton,
    ustonn_prop = tonn_prop,
    usmilex = milex,
    usmilper = milper,
  )  
view(USNaval)

#Importing PolityIV data and creating two seperate datasets to apply to both members of the dyad.
PolityIV <- as.data.frame(read_dta("C:/Users/packe/Documents/Methods_3/Paper/Datasets/PolityIV.dta"))

PolityIV1 <- select(PolityIV, year, ccode, democ, autoc, polity, polity2)
PolityIVa <- PolityIV1 %>%
  rename(
    ccodea = ccode,
    democa = democ,
    autoca = autoc,
    polity2a = polity2
  ) 
view(PolityIVa)

@@

Alliance <- as.data.frame(read_dta("C:/Users/packe/Documents/Methods_3/Paper/Datasets/alliance_v4.1_by_directed_yearly.dta"))
Alliance <- Alliance %>%
  rename(
    ccodea = ccode1,
    ccodeb = ccode2, 
  )

@@

politically_relevant <- as.data.frame(read_dta("C:/Users/packe/Documents/Methods_3/Paper/Datasets/politically_relevant.dta"))
view(politically_relevant)

@@

##Introduction

@@

##Review of Literature

@@

##Theory
	In this paper I make several assumptions about the international system as well as the actors who occupy it. First, unlike John Mearsheimer I assume that states are able to become a global hegemon, and they are able to do so thanks to an overwhelming advantage in naval capabilities. Naval power allows a state to control the international shipping lanes, land troops on foreign shores, and protect supply routes from across the sea thus allowing invasion forces to remain supplied and reinforced. Without a powerful navy, no state can become a hegemon, but this does not mean that a state cannot gain the naval capabilities necessary to create and enforce and international order.
	Other states in the international system will understand that the hegemon has an interest in maintaining the status quo and that the hegemon will punish those who try to upset the international order. With this in mind, states will be less inclined to engage in conflictual behavior and risk the hegemon responding. Unless that state shares a land border with the hegemon, thill will mean that the state will be wary of the hegemon’s naval capabilities as it will be their navy that allows them to respond to any action in the international system. In this paper I present two simple hypotheses:
H1) Greater naval capability by the hegemon will result in a lower propensity for a dyad to experience a MID in a given year.
H2) Greater naval capability by the hegemon will result in a lower propensity for a dyad to experience a MID which involves the use of force in a given year.
I expect the results of my statistical test to show a statistically significant negative relationship between the hegemon’s naval capabilities and the likelihood of a MID occurring. Additionally, I expect that each of the additional controls I am including will be statistically significant, as each of them have been found to be so in different studies.

##Method
	To test my hypothesis that increases in the naval capabilities of a hegemonic power can reduce the propensity of states to engage in militarized disputes or violent militarized disputes, I use the time-tested method of logistic regression. This test observes and records the effects of the independent variables and controls on a binary dependent variable. One of the benefits of logistic regression is that we can get the odds ratio of a particular outcome occurring, allowing us to conclude how much more likely the event is to occur based on increases in the independent variables. Two dependent variables will be tested in this paper, the first being the occurrence of a MID in a given dyad-year. The second is the occurrence of a violent MID in a given dyad year defined asa MID which has achieved a hostility level of “use of force” or “war.” Bivariate logistic models will be tested looking only at the dependent and independent variables. This will be followed by larger models which will include a series of state and dyad level controls. These models will also be replicated using the more traditional measurement of total tonnage of the U.S. fleet as a robustness check.

##Data
The data for this project is drawn from several sources. Data on militarized disputes is drawn from the Militarized Interstate Dispute (MIDs) dataset produced by the correlates of war project. This data includes information of militarized disputes dating as far back as the early 1800s and as recently as 2012. The data includes the start and end dates of the dispute, which state was involved in the dispute coded by both country code and country abbreviation, and who the initiator of the dispute was at the time it started. Also included in this dataset is a variable which identifies the highest level of hostility achieved in this dispute. This is a catagorical variable with observations for threats of using force, displays of force, use of force, or interstate war. An example of threat to use force against an opponent would be former President George W. Bushes televised ultimatum to Saddam Hussein where he offered the former Iraqi leader an amount of time to abdicate his power, or else the US would remove him by force. Displays of force can include the deploying of troops to a contested region such as the recent deployment of the aircraft carrier USS Abraham Lincoln to the Persian Gulf, or the sailing of U.S. warships through the Taiwan Strait. The use of force would be any action in which military forces are used to attack a specified target, but the two countries are not yet at a level of open warfare. The U.S. missile strikes on Syrian airfields or the Israeli bombing of the Iraqi powerplant would constitute a use of force short of open warfare. The highest level of hostility is the case of interstate war such as the World Wars. This variable will be used and modified in this research into two separate deponent variables which are discussed in more detail below.

@@

##Variables
There are two dependent variables that are tested in this paper, the occurrence of a MID in a dyad year and the occurrence of a MID which involved either the use of force or resulted in an interstate war. Both variables are binary and

@@

##Analysis
Figure four gives the unstandardized coefficents of the bivariate test of my independant and dependant variables. Total US tonnage, the traditional measure used by a vast majority of the naval capabilities literature, appears to demonstrate a positive relationship on the occurence of a militaized dispute. However, the measure for naval capabilities which takes into account the quality of fleet shows a statistically significant negative relationship.

@@

##Conclusion
In this paper I have sought to form a more complete measurement for the naval capability of a hegemonic power that went beyond the traditional measurements of fleet size. By factoring in the level of capital intensity per capita of the state in its military I have created a measure which incorporates the traditional measurement for systemic tonnage as well as taking into consideration the quality of the state’s military. The result was a variable which demonstrated a monotonic increase in naval capabilities from the early 1950s to the present day. This stands in contrast to the traditional measurement of total naval tonnage which demonstrated a near monotonic decrease across the same time period, despite substantial improvements in the quality of the fleet.  
After creating my new variables and running my statistical tests, I have found evidence to support both of my hypotheses. The logistic regressions demonstrate that there is a statistically significant negative relationship between the naval capability of the hegemon and the propensity for dyads to experience a militarized dispute. This research has plenty of room for improvement in future iterations. First, this paper has only examined the impact of naval capabilities during the time of a single historic hegemon. Future research may include observations for the naval capabilities of past super powers such as the United Kingdom, Spain, and the Netherlands, three hegemonic powers which possessed substantial naval capabilities. New and improved measurements for naval capabilities as more data becomes available as well as the inclusion of additional control parameters are sure to be tested as well.

>>>>

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Global Reach or Out of Reach"
author: 
- name: Addison Huygens
  affiliation: University of Iowa
abstract: "Research which has examined the effects of naval capabilities on conflict behavior have done so by examining different measures for the size of a state's navy. However, this does not capture the actualy capabilities of a country's fleet. In this article I create a new measurement for naval capability and use it to test the effect of a hegemon's naval power on the conflict behavior of other states in the international system. I find that higher levels of naval power held by the hegemon reduces the liklihood that a militarized dispute will occur."
keywords: "Naval power hegemony"
date: "May 10, 2019"
fontsize: 12pt
spacing: single
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), "Naval\\.bib$"))`}
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
---

@@

dyadic_mid_3_1_may_2018 <- read_dta("Data/dyadic_mid_3.1_may_2018.dta")

@@

#Relative capabilities data
NMC_5_0 <- read_dta("Data/NMC_5_0.dta")
NMC <- select(NMC_5_0, ccode, year, irst, milex, milper, pec, tpop, upop, cinc)
NMCA <- NMC %>%
  rename(
    ccodea = ccode,
    irsta = irst,
    milexa = milex,
    milpera = milper,
    peca = pec,
    tpopa = tpop,
    upopa = upop,
    cinca = cinc
  ) 
NMCA

@@

#Importing naval data, removing all but the data for the U.S., keeping only relevant variables.
crisher_souva_navaldata_II <- read_dta("Data/crisher_souva_navaldata_II.dta")
Naval1 <- filter(crisher_souva_navaldata_II, cowcode == 2)
Naval2 <- select(Naval1, year, totton, tonn_prop, milex, milper)
USNaval <- Naval2 %>%
  rename(
    ustoton = totton,
    ustonn_prop = tonn_prop,
    usmilex = milex,
    usmilper = milper,
  )

#Importing PolityIV data and creating two seperate datasets to apply to both members of the dyad.
PolityIV <- read_dta("Data/PolityIV.dta")

PolityIV1 <- select(PolityIV, year, ccode, democ, autoc, polity, polity2)
PolityIVa <- PolityIV1 %>%
  rename(
    ccodea = ccode,
    democa = democ,
    autoca = autoc,
    polity2a = polity2
  )

@@

Alliance <- read_dta("Data/alliance_v4.1_by_directed_yearly.dta")
Alliance <- Alliance %>%
  rename(
    ccodea = ccode1,
    ccodeb = ccode2, 
  )

@@

politically_relevant <- read_dta("Data/politically_relevant.dta")
view(politically_relevant)

@@

## Introduction

@@

## Review of Literature

@@

## Theory
	In this paper I make several assumptions about the international system as well as the actors who occupy it. First, unlike John Mearsheimer I assume that states are able to become a global hegemon, and they are able to do so thanks to an overwhelming advantage in naval capabilities. Naval power allows a state to control the international shipping lanes, land troops on foreign shores, and protect supply routes from across the sea thus allowing invasion forces to remain supplied and reinforced. Without a powerful navy, no state can become a hegemon, but this does not mean that a state cannot gain the naval capabilities necessary to create and enforce and international order.
	Other states in the international system will understand that the hegemon has an interest in maintaining the status quo and that the hegemon will punish those who try to upset the international order. With this in mind, states will be less inclined to engage in conflictual behavior and risk the hegemon responding. Unless that state shares a land border with the hegemon, thill will mean that the state will be wary of the hegemon’s naval capabilities as it will be their navy that allows them to respond to any action in the international system. In this paper I present two simple hypotheses:
H1) Greater naval capability by the hegemon will result in a lower propensity for a dyad to experience a MID in a given year.
H2) Greater naval capability by the hegemon will result in a lower propensity for a dyad to experience a MID which involves the use of force in a given year.
I expect the results of my statistical test to show a statistically significant negative relationship between the hegemon’s naval capabilities and the likelihood of a MID occurring. Additionally, I expect that each of the additional controls I am including will be statistically significant, as each of them have been found to be so in different studies.

## Method
	To test my hypothesis that increases in the naval capabilities of a hegemonic power can reduce the propensity of states to engage in militarized disputes or violent militarized disputes, I use the time-tested method of logistic regression. This test observes and records the effects of the independent variables and controls on a binary dependent variable. One of the benefits of logistic regression is that we can get the odds ratio of a particular outcome occurring, allowing us to conclude how much more likely the event is to occur based on increases in the independent variables. Two dependent variables will be tested in this paper, the first being the occurrence of a MID in a given dyad-year. The second is the occurrence of a violent MID in a given dyad year defined asa MID which has achieved a hostility level of “use of force” or “war.” Bivariate logistic models will be tested looking only at the dependent and independent variables. This will be followed by larger models which will include a series of state and dyad level controls. These models will also be replicated using the more traditional measurement of total tonnage of the U.S. fleet as a robustness check.

## Data
The data for this project is drawn from several sources. Data on militarized disputes is drawn from the Militarized Interstate Dispute (MIDs) dataset produced by the correlates of war project. This data includes information of militarized disputes dating as far back as the early 1800s and as recently as 2012. The data includes the start and end dates of the dispute, which state was involved in the dispute coded by both country code and country abbreviation, and who the initiator of the dispute was at the time it started. Also included in this dataset is a variable which identifies the highest level of hostility achieved in this dispute. This is a catagorical variable with observations for threats of using force, displays of force, use of force, or interstate war. An example of threat to use force against an opponent would be former President George W. Bushes televised ultimatum to Saddam Hussein where he offered the former Iraqi leader an amount of time to abdicate his power, or else the US would remove him by force. Displays of force can include the deploying of troops to a contested region such as the recent deployment of the aircraft carrier USS Abraham Lincoln to the Persian Gulf, or the sailing of U.S. warships through the Taiwan Strait. The use of force would be any action in which military forces are used to attack a specified target, but the two countries are not yet at a level of open warfare. The U.S. missile strikes on Syrian airfields or the Israeli bombing of the Iraqi powerplant would constitute a use of force short of open warfare. The highest level of hostility is the case of interstate war such as the World Wars. This variable will be used and modified in this research into two separate deponent variables which are discussed in more detail below.

@@

## Variables
There are two dependent variables that are tested in this paper, the occurrence of a MID in a dyad year and the occurrence of a MID which involved either the use of force or resulted in an interstate war. Both variables are binary and

@@

## Analysis
Figure four gives the unstandardized coefficents of the bivariate test of my independant and dependant variables. Total US tonnage, the traditional measure used by a vast majority of the naval capabilities literature, appears to demonstrate a positive relationship on the occurence of a militaized dispute. However, the measure for naval capabilities which takes into account the quality of fleet shows a statistically significant negative relationship.

@@

## Conclusion
In this paper I have sought to form a more complete measurement for the naval capability of a hegemonic power that went beyond the traditional measurements of fleet size. By factoring in the level of capital intensity per capita of the state in its military I have created a measure which incorporates the traditional measurement for systemic tonnage as well as taking into consideration the quality of the state’s military. The result was a variable which demonstrated a monotonic increase in naval capabilities from the early 1950s to the present day. This stands in contrast to the traditional measurement of total naval tonnage which demonstrated a near monotonic decrease across the same time period, despite substantial improvements in the quality of the fleet.  
After creating my new variables and running my statistical tests, I have found evidence to support both of my hypotheses. The logistic regressions demonstrate that there is a statistically significant negative relationship between the naval capability of the hegemon and the propensity for dyads to experience a militarized dispute. This research has plenty of room for improvement in future iterations. First, this paper has only examined the impact of naval capabilities during the time of a single historic hegemon. Future research may include observations for the naval capabilities of past super powers such as the United Kingdom, Spain, and the Netherlands, three hegemonic powers which possessed substantial naval capabilities. New and improved measurements for naval capabilities as more data becomes available as well as the inclusion of additional control parameters are sure to be tested as well.
